Toward theories of deliberated preferences

### 2020-10-30 simple.tex
Improved pres. normative; started better pres. descriptive

As a result, normative approaches are difficult to validate: being considered as non-descriptive in their goal, normative claims cannot be submitted to empirical tests that would permit to verify or falsify them.
Instead, the validity of a normative approach is to be determined by examining the convincingness power of the normative axioms on which it rests.
This validity criterion is unsatisfactory, as it is unable to settle disputes between experts when they arise. In supplement, agreement between experts in decision theory is arguably not enough to ground the practical applicability of some normative axioms: it is desirable to check that decision makers to which the prescription is addressed agree with the particular instanciation of these axioms in the specific decision problem under study. But no systematic, formally defined procedure has been proposed to proceed with this check, to the best of my knowledge. 
In practical applications, an expert may informally discuss the axioms with the decision maker in order to check endorsement, either in general, or by discussing violations on a case by case basis when they are observed during an interactive elicitation session. 
The resulting state of conviction arguably depends in part on the rethorical ability of the expert, not solely on the normative strength of the axioms. It may lead to a decision maker accepting a decision procedure considered unsound by many experts in the field, such as AHP. 
This problem is compounded by the enormous constraining power gained when axioms interact, as the celebrated theorem of Arrow shows dramatically: axioms that appear innocuous when taken alone may precipitate undesirable results when combined.

By contrast, this difficulty of validation is much alleviated with descriptive approaches. Such approaches can be validated by comfronting their predictions to empirical data. When two descriptive theories disagree on their predictions, it is often possible to show that at least one of them is wrong with a suitable experiment. 
Alàs, descriptive approaches as currently existing in decision theory are not suitable to obtain valid prescriptions.

@@

It is by no means my intent to claim that descriptive approaches never face validation difficulties. Experimental tests to arbitrate between theories may not be feasible in practice, descriptions are not always solely about the future, falsificationist approaches in turn raise epistemological difficulties, and so on. Correspondingly, the possible confrontation to empirical data of the theories defined here will likely not suffice to close every discussion about validation. My less naïve claim, that should suffice to motivate the present endeavor, is that it is often a good thing for a theory to make empirically testable claims, from the point of view of the possibilities of solving disputes and validating theories.

As the plural indicates, this article defines a kind of theories, called theories of \ac{DJ}, not a specific theory of \ac{DJ}. Specific theories of \ac{DJ} are tailored to a set of situations and yield claims relevant to those situations. For example, a specific theory of \ac{DJ} may be tailored to situations of choice between lotteries involving money. This article adopts a more general point of view. It proposes a definition of what theories of \ac{DJ} should look like, and proves claims that hold for any specific theory of \ac{DJ}, showing that such theories are indeed normative and descriptive.

@@

These definitions avoid to define the concept of theory itself, for defining precisely such a fundamental concept raises serious difficulties. Is Prospect theory in general a theory, or only some of its variants, sufficiently precisely defined? Can a theory be extremely specific, or is some level of generality required to deserve being called a theory? Must a theory consist in a non-trivial assembly of laws, some deducible from others? It is not necessary for the purpose of the present discussion to take position on these issues, rather, the focus is on the distinction between normative and descriptive approaches. An intuitive understanding of the concept of theory will suffice.
Similarly, I sidestep insuperable difficulties pertaining to capturing in a short definition the concept of a descriptive theory in general: some descriptive theories in a more general sense may propose to describe without predicting, for example (historical decisions or trends come to mind); such endeavor simply does not enter the scope of this discussion. For similar reasons, I do not claim that the term normative is unanimously used in the literature as defined here. This definition encompasses theories used in approaches that do not necessarily self-describe as normative \citep{bell_decision_1988, meinard_rationality_2019}; conversely, some theories considered as normative in the literature, such as some social choice theories, may not fall into the definition proposed here, depending on how they are interpreted.

@@

Another contrast can now be introduced: the contrast between shallow and reflective behavior. \emph{Shallow behavior} is defined as the everyday, usual behavior of an individual. 
This is the behavior whose observation defines what \citet[p.\ 16]{von_neumann_theory_2004} describe as the “intuitive” preferences of an individual, those originating from an “immediate sensation” (cited in \citet{fishburn_retrospective_1989}).
\emph{Reflective behavior}, by contrast, is defined as the behavior that enacts informed, pondered decisions, those that would be taken after due consideration.
% an individual has consciously adopted after having pondered everything that is possibly relevant to the decision. 
%This pair of definitions is, again, vague and unsatisfactory for a deep philosophical discussion, but hopefully sufficient for the present one. 
%What is meant by “pondering about everything that is possibly relevant to the decision” will be defined more precisely later; an intuitive understanding of this phrase will do at this stage.
This distinction draws on our capacity, as human beings, to think about our shallow preferences and consider some of them as mistaken \citep{frankfurt_freedom_1971}. 
A famous example of such a situation is the one of \citet[pp.\ 101--103]{savage_foundations_1972} having realized, after due consideration, that his initial declared preference in a decision problem presented by \citet{allais_so-called_1979} was mistaken.

@@

Let me state how the approach proposed here to capture reflective preferences differs from those proposed in the literature.

The review of existing approaches aiming at obtaining a model of the reflective preferences of an individual distinguishes two kinds of approaches. \Cref{sec:purification} mentions formal techniques, that can be applied by following an explicit procedure, but that do not consider the reaction of individuals to arguments; while \cref{sec:informal} reviews studies that involve informal discussions with individuals, which do consider arguments, but require involving experts. By contrast, this article presents theories that can be confronted to empirical data in a formal, systematic way, and that consider the reactions of individuals to arguments.

@@

@@

These proposals leave unspecified how to confirm or disconfirm them empirically, thus, in particular, how to empirically investigate disagreements between theories: if two theories, on the basis of the same observed data, postulate that they correspond to different purified preferences, the disagreement may at best be arbitrated using ad-hoc, case by case criteria to assess for their relative plausibility. By contrast, the present proposal indicates a clear, systematic empirical way to abitrate disagreements.
%These proposals deserve a detailed comparison to the current one, but this requires a treatment that cannot be given in this article. Let me simply highlight one major difference. which can be done with a high level description of these approaches using a variant of the framework that \citet{rubinstein_eliciting_2011} describe.
%Consider a set of possible observations $S$, describing shallow behavior, and a set of possible “rational” preference relations $R$. The set $S$ may contain observations of choice instances, for example, given the bundle $B$, the individual chose $b \in B$; or may contain all possible linear orders over some set of alternatives. The set $R$ contains all relations that satisfy some set of axioms of rationality, representing what the modeler considers as an acceptably purified relation. \Citet{bleichrodt_making_2001}, for example, use for $R$ the preference relations satisfying \ac{SEU}.

@@

It also remains to be clarified how to check on a principled, systematic basis, whether all relevant arguments have been presented faithfully. Comfronted with 
%(possibly self-designated \commentYM{ceci est un tout autre problème, dont la mention ici brouille le propos}) 
experts with enough assertiveness, a decision-maker can be convinced to follow principles of decision-making that several experts in the field of decision-theory would consider incorrect. 
%As an illustration, after discussing the weaknesses of the \ac{AHP}, \citet[p.\ 53]{howard_foundations_2007} write \commentYM{s}: “Why, then, do inferior processes find favor with decision makers? The answer is that they do not force you to think very hard or to think in new ways.” \commentYM{je ne vois pas en quoi ceci est censé illustrer le propos précédent (ou alors je n'ai rien compris à la citation). J'ai l'impression que la citation dit que, parfois, il y a des forces psychologiques qui s'opposent à l'adoption d'un processus de décision, quand bien même celui-ci aurait des qualités indéniables}
Beyond the controversial \ac{AHP} mentioned in the introduction, famous decision analysts have voiced disagreements with normative theories of decision making \citep{ellsberg_risk_1961, allais_so-called_1979} that are however generally applied in the field. \commentYM{il me semble ici intenable de faire l'économie de définir ce que tu appelles "normative theories"} 
\Citet[p.\ 669]{ellsberg_risk_1961} concluded his influential article as follows: “[In some situations, for some people,] the Bayesian or Savage approach gives [...] bad advice. [Some people] act in conflict with the axioms deliberately, without apology, because it seems to them the sensible way to behave. Are they clearly mistaken?” \commentYM{d'une manière générale, dans toute la section 1, l'avalanche de citation me semble nuire au propos. Cela induit des ruptures dans le style, et force le lecteur à se pencher sur des questions d'interprétation des citations qui l'empêchent de se focaliser sur ton propos à toi}
This rethorical question illustrates the need of systematic procedures able to determine whether decision-makers, when confronted with the relevant arguments, accept the soundness of the principles that the recommendation rests on (this has been somewhat explored, but not systematically, for lack of a systematic way of confronting decision makers to arguments \citep{slovic_who_1974, stanovich_discrepancies_1999}).

@@

\commentOCf{J’ai un peu réduit les citations, j’espère que ça va mieux. Concernant celle-ci en particulier, elle me semble très utile. Je trouve stupéfiant qu’on puisse trouver ces propos dans la littérature, écrits noir sur blanc. Si je ne les cite pas, un lecteur pourrait (ou devrait) ne pas me croire quand j’affirme que des gens trouvent que ce n’est pas grave que deux études soi-disant scientifiques arrivent à des conclusions contradictoires. La citation étendue permet de comprendre, je crois, leur point de vue, et de se faire une idée de ce qu’ils veulent dire par là.}

@@

\commentOCf{Je propose de trancher entre approches normatives en utilisant des données empiriques, plutôt que simplement la force de conviction des axiomes (qui parlent différemment à différentes personnes et peuvent conserver les désaccords intacts). Je ne sais pas si c’est plus clair maintenant.}

@@

First, the previous proposal requires the individual to declaratively take position on notions that may be unclear to him, as they include some possibly counterfactual situations, and as the concept of turning an argument into an uneffective argument may not be understood in the same way by everybody. 
In order for the conclusions about $i$’s \ac{DJ} to hold, we must be sure that $i$ understands the semantics of the relations on which it is based as we do. It is therefore desirable that these notions have unambiguous semantics. It is undoubtedly an important strength of \citet{samuelson_foundations_1983}, \citet{von_neumann_theory_2004} or \citet{savage_foundations_1972}’s theories that they use as primitive notions observable entities whose semantics is unambiguous. This is not to say that these proposals cannot be criticized. For example, the claim that the notion of preference can be disposed of entirely has been much criticized \citep{hausman_preference_2011, wong_foundations_2006}, with good reasons; and observations of shallow behavior, though having a clear semantics, are insufficient to deduce normative claims, as has been amply remarked and as argued in this very article. This is also not to say that declarations should never be used and that only choice acts should matter to any theory in economics.
Nonetheless, as \citet[p.\ 16]{von_neumann_theory_2004} aptly summarize, “It is clear that every measurement – or rather every claim of measurability – must ultimately be based on some immediate sensation, which possibly cannot and certainly need not be analyzed any futher”. A weakness of the previous proposal for studying \acp{DJ} is that the observable relations it uses may be considered inadequate primitive notions. This article improves on this aspect by postulating a non-ambiguous observable relation, that does not require $i$ to take position declaratively on relative qualities of arguments; but rather to choose (by declaration of preference or by act) among bundles, or to declare whether $i$ adheres to a given proposition in her current perspective. (In the latter case, the clarity of the semantics of the observable relation will depend on the clarity of the propositions.)

@@

Finally, the current proposal links \ac{DJ} and deliberated preferences. \commentYM{indispensable de dire explicitement quelle est la différence entre les deux dès maintenant}

@@

Such descriptive theories can take multiple forms, and some of them would permit to see clearly that the theory is descriptive in the required sense, that is, that its claims can be falsified given observational data. 
But the task we face, for such theories to be useful, is more complex than this minimal requirement: we have to find forms of theories that are not hugely demanding on the part of the theorist. In particular, we do not want to oblige the theorist to predict the whole of the argumentative attitude of individuals if this is not required to obtain information about her \ac{DJ}. Thus, \cref{sec:descriptive} defines descriptive theories of \ac{DJ} that only claim to be able to answer counter-arguments, in a sense that will be made precise. The desire that its claim be falsifiable is also defined formally. More formally, the result that is obtained in that section, and that is the central result of this article, is that if a theory is convincing and normatively adequate, then it is valid. That a theory is convincing is (as also shown in the same section) the same as demanding that it resists all empirical falsification attempts. It is normatively adequate if it satisfies the normative axiom described here above informally (and formally in \cref{sec:normative}). And it is valid if the propositions that it claims are in someone’s \ac{DJ} are indeed in that person’s \ac{DJ}. This central result thus amounts to prove the main claim of this article, namely, that accepting a weak normative axiom, supplemented with empirical claims, is enough to capture someone’s reflective preferences.

A possible objection to the usefulness of this approach should perhaps be lifted before delving into the formal descriptions of this rough sketch.
It may be argued that because this proposition needs a normative axiom, which cannot be validated empirically, it does not fare better than the alternative propositions examined in the previous sections. \ac{SEU}, for example, also claim to obtain reflective preferences on the basis of normative axioms.
I grant that it is impossible to get completely rid of normative axioms, because prescriptions must be linked to descriptions, and this cannot be done in a complete empirical way.
However, it is good to lighten as much as possible the content of normative axioms, and defer as much content as possible to descriptive claims. The descriptive parts of theories can be confronted to observations, a natural way to lift disagreements that has often proved useful in the history of science (again, accepting this claim does not require to ignore that observations themselves rest upon theories). Arguably, the normative axiom proposed here is much weaker than those of a theory such as \ac{SEU}, in the sense of having less substance or of being compatible with more preference relations. (This claim can only be valid at an intuitive level: because \ac{SEU} does not use arguments, its axioms cannot be compared formally to mine.) This does not prevent theories of \ac{DJ} to propose substantive, non-trivial claims about reflective preferences. It simply moves the debate from the world of discussions between experts about the validity of the axioms to the world of observations of acceptability of arguments by the persons to which these theories claim to be applicable.
It is also worth noting that this axiom has a specific form that may permit decision makers to reconstruct the reasoning proposed by the theory so as to be able to check that they apply to their situation. Detailed exploration of this aspect is deferred to future work.

@@

\section{Informal presentation}
\commentOCf{Je pense changer d’approche : enlever cette section et intégrer les explications informelles directement dans la présentation formelle. À faire…}
\commentYM{Le lecteur n'a aucune idée de ce qu'il doit s'attendre à trouver dans cette section, et de la mesure dans laquelle ça va permettre d'avancer dans la résolution de la problématique : lecteur perdu = pas bien}
\commentYM{par qui ? dans le cadre de l'analyse ou en amont ? et pourquoi sur la seule base de la soundness de l'axiome ? pourquoi pas sur la base de ses implications dans des cas concrets ?}
An empirical theory for a given normative theory, by contrast, makes empirical claims, hence, can be confronted to observations \commentYM{quelle différence avec les implications empiriques d'une théorie normative ?}. What I call here an empirical theory, for simplicity, is not however purely empirical, but can also be considered normative, for two reasons. First, its conclusions inevitably rest on the normative validity of the normative theory for which it is built. Second, it claims to capture preferences that provide an appropriate basis for recommendation. Empirical theories of \ac{DJ} could thus be called empirical normative theories of \ac{DJ}.\commentYM{est-ce la même chose que ce qui est déjà appelé à peu près comme ça dans la littérature ? Je parle par exemple de toutes les théories empiriques de la justice}

\subsection{Normative theories}
Normative theories define an agenda: a set of propositions that we are interested in knowing the deliberated judgment of individuals about ; a set of individuals that the theory claims applies to; a set of arguments;\commentYM{pourquoi est-ce que cela fait partie de la théorie ? c'est plutôt un des éléments du contexte d'une de SES APPLICATIONS à mon sens. L'expression  "normative model" serait selon moi bien plus appropriée si on veut que les détails du cas d'application fassent partie de l'objet. "Theory" a de très fortes connotations de généralité} and a protocol of observation of the judgments an individual adopts towards a proposition after exposure to various arguments \commentYM{pourquoi le protocole d'obs est-il déjà évoqué ici ? j'aurais cru que ce protocole relève de la théorie empirique qui se greffe sur la théorie normative en question. Par ailleurs, je croyais que c'était l'axiome de la théorie normative qui devait être mis à l'épreuve, pas une instantiation sur une propositon. De l'axiome, il n'est plus question. Où est-il ?}.\commentYM{il manque dans la "theorie normative" ce que j'aurais considéré comme le plus important : un modèle de la force de conviction des arguments et de leurs liens avec les propositions. J'aurais dit que l'axiome de la "théorie" se cacherait là-dedans}

Examples of propositions to be included in \commentYM{the}agenda \commentYM{might}include: “I ought to pay at most \$100 for purchasing bundle $b$”, “I ought to acquire alternative $a$ rather than $b$ if it is available in the same conditions”; “I ought not to perform action $a$ because it is immoral”.

The set of individuals in a normative theory represents the situations, socio-economic conditions, acquired diplomas, or other observable features of individuals that must be satisfied for the theory to apply to them \commentYM{cela ressemble plus à un domaine d'application, qu'à un ensemble d'individus auxquels ça s'applique. Dans ce cas pourquoi ne pas le dire clairement ?}. (More about this in conclusion. \TODO{})

Arguments are defined here as anything that can be presented to an individual that may possibly influence her judgment. As an example, an argument can be a text presenting the features of an alternative \commentYM{donc un texte qui ne prend pas de position sur l'alternative, qui est purement descriptif, est un argument ? cela me semble raisonnable, mais ça mérite discussion}, or arguing for the (im)morality of an action. The term argument evokes a debate, a metaphor appropriate for the comparison of theories that will be described later; but that should be taken with a grain of salt: an image, a tag price, a poem can be an argument in this article.

>>>>

As a result, normative approaches are difficult to validate: being considered as non-descriptive in their goal, normative claims cannot be submitted to empirical tests that would permit to verify or falsify them.
Instead, the validity of a normative approach is to be determined by examining the convincingness power of the normative axioms on which it rests.
\commentYM{par qui ? dans le cadre de l'analyse ou en amont ? et pourquoi sur la seule base de la soundness de l'axiome ? pourquoi pas sur la base de ses implications dans des cas concrets ?}
This validity criterion is unsatisfactory, as it is unable to settle disputes between experts when they arise. In supplement, agreement between experts in decision theory is arguably not enough to ground the practical applicability of some normative axioms: it is desirable to check that decision makers to which the prescription is addressed agree with the particular instanciation of these axioms in the specific decision problem under study. But no systematic, formally defined procedure has been proposed to proceed with this check, to the best of my knowledge. 
In practical applications, an expert may informally discuss the axioms with the decision maker in order to check endorsement, either in general, or by discussing violations on a case by case basis when they are observed during an interactive elicitation session. 
The resulting state of conviction arguably depends in part on the rethorical ability of the expert, not solely on the normative strength of the axioms. It may lead to a decision maker accepting a decision procedure considered unsound by many experts in the field, such as the \ac{AHP} \citep{smith_anniversary_2004}. 
This problem is compounded by the enormous constraining power gained when axioms interact, as the celebrated theorem of Arrow shows dramatically: axioms that appear innocuous when taken alone may precipitate undesirable results when combined.

By contrast, this difficulty of validation is much alleviated with descriptive approaches. Such approaches can be validated by comfronting their predictions to empirical data. When two descriptive theories disagree on their predictions, it is often possible to show that at least one of them is wrong with a suitable experiment. 
\commentYM{quelle différence avec les implications empiriques d'une théorie normative ?}
Alàs, descriptive approaches as currently existing in decision theory are not suitable to obtain valid prescriptions.

@@

It is by no means my intent to claim that descriptive approaches never face validation difficulties. Experimental tests to arbitrate between theories may not be feasible in practice, descriptions are not always solely about the future, falsificationist approaches in turn raise epistemological difficulties, and so on. Correspondingly, the possible confrontation to empirical data of the theories defined here is no magical means to definitively solve every difficulty about validation. My less naïve claim, that should suffice to motivate the present endeavor, is that it is often a good thing for a theory to make empirically testable claims, from the point of view of the possibilities of solving disputes and validating theories.

As the plural indicates, this article defines a kind of theory, namely, theories of \ac{DJ}, not a specific theory of \ac{DJ}. Specific theories of \ac{DJ} are tailored to a set of situations and yield claims relevant to those situations. For example, a specific theory of \ac{DJ} may be tailored to situations of choice between lotteries involving money. This article adopts a more general point of view. It proposes a definition of what theories of \ac{DJ} should look like, and proves claims that hold for any specific theory of \ac{DJ}, showing that such theories are indeed normative and descriptive.

@@

These definitions avoid to define the concept of theory itself, for defining precisely such a fundamental concept raises serious difficulties. Is Prospect theory in general a theory, or only some of its variants, sufficiently precisely defined? Can a theory be extremely specific, or is some level of generality required to deserve being called a theory? Must a theory consist in a non-trivial assembly of laws, some deducible from others? It is not necessary for the purpose of the present discussion to take position on these issues, rather, the focus is on the distinction between normative and descriptive approaches. An intuitive understanding of the concept of theory will suffice.
Similarly, I sidestep insuperable difficulties pertaining to capturing in a short definition the concept of a descriptive theory in general: some descriptive theories in a more general sense may propose to describe without predicting, for example (historical decisions or trends come to mind); such endeavor simply does not enter the scope of this discussion. For similar reasons, I do not claim that the term normative is unanimously used in the literature as defined here. The present definition encompasses theories used in approaches that do not necessarily self-describe as normative \citep{bell_decision_1988, meinard_rationality_2019}; conversely, some theories considered as normative in the literature, such as some social choice theories, may not fall into the definition proposed here, depending on how they are interpreted.

@@

Another contrast can now be introduced: the contrast between shallow and reflective behavior. \emph{Shallow behavior} is defined as the everyday, usual behavior of an individual. 
This is the behavior whose observation defines what \citet[p.\ 16]{von_neumann_theory_2004} describe as the “intuitive” preferences of an individual, those originating from an “immediate sensation” (the quotes from their article used here are cited by \citet{fishburn_retrospective_1989}).
\emph{Reflective behavior}, by contrast, is defined as the behavior that enacts informed, pondered decisions, those that would be taken after due consideration.
% an individual has consciously adopted after having pondered everything that is possibly relevant to the decision. 
%This pair of definitions is, again, vague and unsatisfactory for a deep philosophical discussion, but hopefully sufficient for the present one. 
%What is meant by “pondering about everything that is possibly relevant to the decision” will be defined more precisely later; an intuitive understanding of this phrase will do at this stage.
This distinction draws on our capacity, as human beings, to think about our shallow preferences and consider some of them as mistaken \citep{frankfurt_freedom_1971}. 
A famous example of such a situation is the one of \citet[pp.\ 101--103]{savage_foundations_1972} having realized, after due consideration, that his initial declared preference in a decision problem presented by \citet{allais_so-called_1979} was mistaken.

@@

Such descriptive theories can take multiple forms, and some of them would permit to see clearly that the theory is descriptive in the required sense, that is, that its claims can be falsified given observational data. 
But the task we face, for such theories to be useful, is more complex than this minimal requirement: we have to find forms of theories that are not hugely demanding on the part of the theorist. In particular, we do not want to oblige the theorist to predict the whole of the argumentative attitude of individuals if this is not required to obtain information about her \ac{DJ}. Thus, \cref{sec:descriptive} defines descriptive theories of \ac{DJ} that only claim to be able to answer counter-arguments, in a sense that will be made precise. The desire that its claim be falsifiable is also defined formally. More formally, the result that is obtained in that section, and that is the central result of this article, is that if a theory resists all empirical falsification attempts, then it is valid, assuming acceptance of the normative axiom described here above informally (and formally in \cref{sec:normative}). A theory is valid if the propositions that it claims are in someone’s \ac{DJ} are indeed in that person’s \ac{DJ}. This central result thus amounts to prove the main claim of this article, namely, that accepting a weak normative axiom, supplemented with empirical claims, is enough to capture someone’s reflective preferences.

A possible objection to the usefulness of this approach should perhaps be lifted before delving into the formal descriptions of this rough sketch.
It may be argued that because this proposition needs a normative axiom, which cannot be validated empirically, it does not fare better than the alternative propositions existing in the literature. \ac{SEU}, for example, also can be considered to claim to obtain reflective preferences on the basis of normative axioms.
I grant that it is impossible to get completely rid of normative axioms, because prescriptions must be linked to descriptions, and this cannot be done in a complete empirical way.
However, it is good to lighten as much as possible the content of normative axioms, and defer as much content as possible to descriptive claims. The descriptive parts of theories can be confronted to observations, a natural way to lift disagreements that has often proved useful in the history of science (again, accepting this claim does not require to ignore that observations themselves rest upon theories). Arguably, the normative axiom proposed here is much weaker than those of a theory such as \ac{SEU}, in the sense of having less substance. (This claim can only be valid at an intuitive level: because \ac{SEU} does not use arguments, its axioms cannot be compared formally to mine.) This does not prevent theories of \ac{DJ} to propose substantive, non-trivial claims about reflective preferences. It simply moves the debate from the world of discussions between experts about the validity of the axioms to the world of observations of acceptability of arguments by the persons to which these theories claim to be applicable.
It is also worth noting that the proposed axiom has a specific form that may permit decision makers to reconstruct the reasoning proposed by the theory so as to be able to check that they apply to their situation. Detailed exploration of this aspect is deferred to future work.

Let us now examine how the approach proposed here to capture reflective preferences differs from those proposed in the literature.

The next two sections review existing approaches aiming at obtaining a model of the reflective preferences of an individual. They distinguish two kinds of approaches. \Cref{sec:purification} mentions formal techniques, that can be applied by following an explicit procedure, but that do not consider the reaction of individuals to arguments; while \cref{sec:informal} reviews studies that involve informal discussions with individuals, which do consider arguments, but require involving experts. By contrast, this article presents theories that can be confronted to empirical data in a formal, systematic way, and that consider the reactions of individuals to arguments.

@@

The field of Empirical social choice \citep{gaertner_empirical_2012} investigates individuals’s judgments about fairness: in a typical laboratory experiment, subjects are described some setting requiring to distribute resources among a set of individuals, and are asked to choose a distribution that they consider fair. These interesting approaches permit to study empirically the sense of fairness of real individuals, and compare them to social choice theories. They are distinct from the currently proposed approach as they do not involve arguments; and do not specifically aim to prescribe to an individual.

@@

These proposals leave unspecified how to confirm or disconfirm them empirically, thus, in particular, how to empirically investigate disagreements between theories: if two theories, on the basis of the same observed data, postulate that they correspond to different purified preferences, the disagreement may at best be arbitrated using ad-hoc, case by case criteria to assess for their relative plausibility. By contrast, the present proposal indicates a clear, systematic empirical way to arbitrate disagreements.
%These proposals deserve a detailed comparison to the current one, but this requires a treatment that cannot be given in this article. Let me simply highlight one major difference. which can be done with a high level description of these approaches using a variant of the framework that \citet{rubinstein_eliciting_2011} describe.
%Consider a set of possible observations $S$, describing shallow behavior, and a set of possible “rational” preference relations $R$. The set $S$ may contain observations of choice instances, for example, given the bundle $B$, the individual chose $b \in B$; or may contain all possible linear orders over some set of alternatives. The set $R$ contains all relations that satisfy some set of axioms of rationality, representing what the modeler considers as an acceptably purified relation. \Citet{bleichrodt_making_2001}, for example, use for $R$ the preference relations satisfying \ac{SEU}.

@@

It also remains to be clarified how to check on a principled, systematic basis, whether all relevant arguments have been presented faithfully. Comfronted with 
%(possibly self-designated \commentYM{ceci est un tout autre problème, dont la mention ici brouille le propos}) 
experts with enough assertiveness, a decision-maker can be convinced to follow principles of decision-making that several experts in the field of decision-theory would consider incorrect. 
%As an illustration, after discussing the weaknesses of the \ac{AHP}, \citet[p.\ 53]{howard_foundations_2007} write \commentYM{s}: “Why, then, do inferior processes find favor with decision makers? The answer is that they do not force you to think very hard or to think in new ways.” \commentYM{je ne vois pas en quoi ceci est censé illustrer le propos précédent (ou alors je n'ai rien compris à la citation). J'ai l'impression que la citation dit que, parfois, il y a des forces psychologiques qui s'opposent à l'adoption d'un processus de décision, quand bien même celui-ci aurait des qualités indéniables}
Beyond the controversial \ac{AHP} mentioned in the introduction, strongly criticized, for example, by \citet[p.\ 53]{howard_foundations_2007}, famous decision analysts have voiced disagreements with normative theories of decision making \citep{ellsberg_risk_1961, allais_so-called_1979} that are however generally applied in the field. \commentYM{il me semble ici intenable de faire l'économie de définir ce que tu appelles "normative theories"} 
\Citet[p.\ 669]{ellsberg_risk_1961} concluded his influential article as follows: “[In some situations, for some people,] the Bayesian or Savage approach gives [...] bad advice. [Some people] act in conflict with the axioms deliberately, without apology, because it seems to them the sensible way to behave. Are they clearly mistaken?” \commentYM{d'une manière générale, dans toute la section 1, l'avalanche de citation me semble nuire au propos. Cela induit des ruptures dans le style, et force le lecteur à se pencher sur des questions d'interprétation des citations qui l'empêchent de se focaliser sur ton propos à toi}
This rethorical question illustrates the need of systematic procedures able to determine whether decision-makers, when confronted with the relevant arguments, accept the soundness of the principles that the recommendation rests on (this has been somewhat explored, but not systematically, for lack of a systematic way of confronting decision makers to arguments \citep{slovic_who_1974, stanovich_discrepancies_1999}).

@@

\commentOCf{J’ai un peu réduit les citations, j’espère que ça va mieux. Concernant celle-ci en particulier, elle me semble très utile. Je trouve stupéfiant qu’on puisse trouver ces propos dans la littérature, écrits noir sur blanc. Si je ne les cite pas, un lecteur pourrait (ou devrait) ne pas me croire quand j’affirme que des gens trouvent que deux études soi-disant scientifiques puissent arriver à des conclusions apparemment contradictoires et être toutes les deux valides. La citation étendue permet de comprendre, je crois, leur point de vue, et de se faire une idée de ce qu’ils veulent dire par là.}

@@

\commentOCf{En gros, je propose de trancher entre approches normatives en utilisant des données empiriques, plutôt que simplement la force de conviction des axiomes (qui parlent différemment à différentes personnes et peuvent conserver les désaccords intacts). Je ne sais pas si c’est plus clair maintenant.}

@@

First, the previous proposal requires the individual to declaratively take position on notions that may be unclear to him, as they include some possibly counterfactual situations, and as the concept of turning an argument into an uneffective argument may not be understood in the same way by everybody. 
In order for the conclusions about $i$’s \ac{DJ} to hold, we must be sure that $i$ understands the semantics of the relations on which it is based as analysts do. It is therefore desirable that these notions have unambiguous semantics. It is undoubtedly an important strength of \citet{samuelson_foundations_1983}, \citet{von_neumann_theory_2004} or \citet{savage_foundations_1972}’s theories that they use as primitive notions observable entities whose semantics is unambiguous. This is not to say that these proposals cannot be criticized. For example, the claim that the notion of preference can be disposed of entirely has been much criticized \citep{hausman_preference_2011, wong_foundations_2006}, with good reasons; and observations of shallow behavior, though having a clear semantics, are insufficient to deduce normative claims, as has been amply remarked and as argued in this very article. This is also not to say that declarations should never be used and that only choice acts should matter to any theory in economics.
Nonetheless, as \citet[p.\ 16]{von_neumann_theory_2004} aptly summarize, “It is clear that every measurement – or rather every claim of measurability – must ultimately be based on some immediate sensation, which possibly cannot and certainly need not be analyzed any futher”. A weakness of the previous proposal for studying \acp{DJ} is that the observable relations it uses may be considered inadequate primitive notions. This article improves on this aspect by postulating a non-ambiguous observable relation, that does not require $i$ to take position declaratively on relative qualities of arguments; but rather to choose (by declaration of preference or by act) among bundles, or to declare whether $i$ adheres to a given proposition in her current perspective. (In the latter case, the clarity of the semantics of the observable relation will depend on the clarity of the propositions.)

@@

Finally, the current proposal links \acp{DJ} and \acp{DP}. \commentYM{indispensable de dire explicitement quelle est la différence entre les deux dès maintenant}

@@

@@

\section{Normative theories}
\label{sec:normative}
\subsection{Definition}
A normative theory of \ac{DJ} is a tuple $(\allprops, \allargs, I, {\allleadsto})$, where $\allprops$ is a set of propositions, $\allargs$ is a set of arguments, $I$ represents a set of individuals, and ${\allleadsto}$ is an observation protocol. Intuitively speaking, the theory claims that the proposed observation protocol, observing the reactions of individuals from $I$ to arguments from $\allargs$, is normatively appropriate to yield recommendations to individuals about the propositions from $\allprops$. These concepts are detailed in this section.
\commentYM{pourquoi le protocole d'obs est-il déjà évoqué ici ? j'aurais cru que ce protocole relève de la théorie empirique qui se greffe sur la théorie normative en question. Par ailleurs, je croyais que c'était l'axiome de la théorie normative qui devait être mis à l'épreuve, pas une instantiation sur une propositon. De l'axiome, il n'est plus question. Où est-il ?}.
\commentYM{il manque dans la "theorie normative" ce que j'aurais considéré comme le plus important : un modèle de la force de conviction des arguments et de leurs liens avec les propositions. J'aurais dit que l'axiome de la "théorie" se cacherait là-dedans}

The set $\allprops$ is called the agenda of the theory. Propositions composing the agenda are not further detailed, they are simply atomic elements. A proposition must be an element such that one can observe whether an individual adheres to it, or whether an individual’s behavior is consistent with it.
\commentYM{pourquoi est-ce que cela fait partie de la théorie ? c'est plutôt un des éléments du contexte d'une de SES APPLICATIONS à mon sens. L'expression  "normative model" serait selon moi bien plus appropriée si on veut que les détails du cas d'application fassent partie de l'objet. "Theory" a de très fortes connotations de généralité} 
Examples of propositions include: “I ought to pay at most \$100 for purchasing bundle $b$”, “I ought to acquire alternative $a$ rather than $b$ if it is available in the same conditions”; “I ought not to perform action $a$ because it is immoral”.

Another primitive elements of a normative theory is the set of arguments $\allargs$.
Arguments are defined here as anything that can be presented to an individual that may possibly influence her judgment. As an example, an argument can be a text presenting the features of an alternative 
\commentYM{donc un texte qui ne prend pas de position sur l'alternative, qui est purement descriptif, est un argument ? cela me semble raisonnable, mais ça mérite discussion}, or arguing for the (im)morality of an action. An argument may be purely descriptive or may forcefully argue in favor of the truth or falsity of one or several propositions.
The term argument generally evokes a debate, a metaphor appropriate for the comparison of theories that will be described later; but that should be taken with a grain of salt: an image, a tag price, a poem can be an argument in this article.

The set of individuals in a normative theory represents the situations, socio-economic conditions, acquired diplomas, or other observable features of individuals that must be satisfied for the theory to apply to them \commentYM{cela ressemble plus à un domaine d'application, qu'à un ensemble d'individus auxquels ça s'applique. Dans ce cas pourquoi ne pas le dire clairement ?}.

The protocol of observation in a normative theory defines how the theory proposes to present a sequence \commentYM{qu'est-ce qu'une séquence ? y a-t-il une notion de temporalité ? de lien logique entre les arguments ? une nécessité que chaque élément de la "séquence" attaque le précédent ? les embranchements sont-ils exclus ? les cycles ? bref, de quoi parles-tu ?}
\commentOCf{C’est le \href{https://en.wikipedia.org/wiki/Sequence}{terme habituel} en mathématiques, à ma connaissance.}
of arguments to an individual, and how to observe her resulting judgment about a proposition of interest. This protocol permits to determine whether a sequence of arguments “leads” a given individual to a given proposition. If the arguments are texts and propositions are of the form “I ought to acquire alternative $a$ rather than $b$”, an example of such a protocol is: present each argument (each text) in the sequence to the individual, asking her to read it carefully; then give the individual a choice between $a$ and $b$; the sequence of arguments is considered to having led the individual to the proposition “I ought to acquire alternative $a$ rather than $b$” iff he chose $a$ rather than $b$.
